Exploring the ``Party City'' effect in the NBA

### 2017-03-29 eda_report.rnw
Added citation. Now including Boston and not Miami as Party City. Examining effect for bottom 50% rank from last season.

\subsection*{References}
``The Benefits of College Athletic Success: An Application of the Propensity Score Design'', by Michael L. Anderson. March, 2017. \href{http://www.mitpressjournals.org/doi/pdfplus/10.1162/REST_a_00589}{paper}

@@

### Define a Party Variable
party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
party.rgx <- paste0('(', party.cities, ')', collapse = "|")

@@

### How often are teams meeting the spread?
hist(data[, mean(outcome == 'W'), by = list(season, team)][, V1],
     breaks = seq(0.3, 0.75, by = 0.025),
     main = "How often are teams meeting the spread?",
     xlab = "Proportion of games for which a team-year meets the spread")
@

@@

@@

<<>>=

@@

tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
            by = list(avg.age = round(avg.age), party, season)][order(avg.age, season, party)]
tmp[order(avg.age, party)]
## ages[, year := as.numeric(year)]
## data[, year := year(date)] # Note quite right.
## data <- merge(data, ages, by = c('team', 'year'), all.x = T)

@@

<<>>=
load(file = 'tmp_data/spreads_with_standings.RData')
model <- glm(outcome == 'W' ~ party + avg.age*party + weekend + ndays.lgame + pct +
                 travel.dist + team + season + team:season,
             family = 'binomial', data = data)

>>>>

\subsection*{References}
``Balancing, Regression, Difference-In-Differences and Synthetic Control Methods: A Synthesis'', by Doudchenko, Imbens. Oct 2016. \href{http://www.nber.org/papers/w22791}{paper} \\
``The Benefits of College Athletic Success: An Application of the Propensity Score Design'', by Michael L. Anderson. March, 2017. \href{http://www.mitpressjournals.org/doi/pdfplus/10.1162/REST_a_00589}{paper}

@@

### Define a Party Variable
#party.cities <- c("miami", "los angeles", "brooklyn", "new york", "atlanta")
party.cities <- c('boston', 'brooklyn', 'los angeles', 'new york')
party.rgx <- paste0('(', party.cities, ')', collapse = "|")

@@

### How often are teams meeting the spread?
hist(data[, mean(outcome == 'W'), by = list(season, team)][, V1],
     breaks = seq(0.3, 0.75, by = 0.025),
     main = "How often are teams meeting the spread?",
     xlab = "Proportion of games for which a team-year meets the spread")

@@

\subsection*{Which teams to include as party cities?}
<<echo=-1>>=

tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
     by = list(last.game.loc, ndays.lgame)][order(last.game.loc, ndays.lgame)]

tmp[ndays.lgame == 1][order(meet.spread)]

<<results='hide'>>=

### What about by game location
ggplot(tmp, aes(x = ndays.lgame, y = meet.spread, label = N)) +
    geom_point() + 
    geom_label(nudge_y = 0.2) +
    geom_hline(yintercept = 0.5) + 
    facet_wrap(facets = ~ last.game.loc)

\subsection*{Does last-season's ranking have an implication?}

<<>>=
tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
            by = list(pct, party)][order(pct)]
ggplot(tmp, aes(x = pct, y = meet.spread, color = as.factor(party), label = N)) + 
    geom_label()
@

@@

<<echo=F>>=

@@

tmp <- data[, list(meet.spread = mean(outcome == 'W'), .N), 
            by = list(avg.age = round(avg.age, 2), party)][
    order(avg.age, party)] %>% na.omit

ggplot(tmp, aes(x = avg.age, y = meet.spread, 
                color = party %>% as.factor, label = N)) + 
    geom_point() +
    geom_label() +
    geom_smooth()
## ages[, year := as.numeric(year)]
## data[, year := year(date)] # Note quite right.
## data <- merge(data, ages, by = c('team', 'year'), all.x = T)

@@

<<>>=
#load(file = 'tmp_data/spreads_with_standings.RData')
model <- glm(outcome == 'W' ~ party + avg.age*party + weekend + ndays.lgame + pct +
                 travel.dist + season,
             family = 'binomial', data = data, subset = pct < 0.5)
